Music Recommendation System Using the Million Song Dataset

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

\begin{abstract}
We attempted to build a recommendation system to recommend songs to users using data from the Million Song Dataset available on Kaggle. The main focus of this paper was to use collaborative filtering algorithms that only utilize user feedback in order to predict what songs users may like. For benchmarking the algorithms, the same Mean Average Precision truncated at 500 recommended songs used for the original Kaggle competition was chosen. It was discovered that probabilistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding a MAP value of 0.04.

@@

\section{Introduction}
With an increasing amount of accessible data, recommendation systems are becoming increasingly viable and important in various industries including retail, entertainment, and even dating. Recommendation systems allow for services to extend their reach to customers by allowing their users to discover things of interest. Hence, there is a strong interest in implementing a good recommendation system so that their user base can grow. It can be said that building such a system is even more important for the music industry because the entire library of artists and songs are far larger compared to something like the movie industry so recommending the right music to a specific user is an even harder task. On the other hand, with the shift from a traditional distribution model utilizing physical CDs or records to digital models where users stream or acquire music from internet services, it is now easier than ever to track user preferences and this presents a new opportunity to better predict and recommend songs to users.

>>>>

\begin{abstract}
We attempted to build a recommendation system to recommend songs to users using data from the Million Song Dataset available on Kaggle. The main focus of this paper was to use collaborative filtering algorithms that only utilize user feedback in order to predict what songs users may like. For benchmarking the algorithms, we used the same Mean Average Precision score truncated at 500 recommended songs that was used in the original Kaggle competition. It was discovered that probabilistic matrix factorization with a MAP value of 0.014 did not improve results much from using a baseline of simply recommending popular songs, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results with the best method giving a MAP value of 0.048.

@@

\section{Introduction}
With an increasing amount of accessible data, recommendation systems are becoming more viable and important in various industries including retail, entertainment, and even dating. Recommendation systems allow for services to extend their reach to customers by allowing their users to discover things of interest. Hence, there is a strong interest in implementing a good recommendation system so that their user base can grow. It can be said that building such a system is even more important for the music industry because the entire library of artists and songs are far larger compared to something like the movie industry so recommending the right music to a specific user is an even harder task. On the other hand, with the shift from a traditional distribution model utilizing physical CDs or records to digital models where users stream or acquire music from internet services, it is now easier than ever to track user preferences and this presents a new opportunity to better predict and recommend songs to users.
